\documentclass[tf-tutorial-all.tex]{subfiles}
\begin{document}

\begin{truefalse}
  The central limit theorem in its approximation form states that for
  rvs $X_i$ with mean $\mu$ and variance $\sigma^2$.
  $$ n^{-1}\sum_{i=1}^n X_i = \bar X_n \stackrel{\cdot}\sim \Norm{\mu, \sigma^2 \sqrt{n}} $$
\begin{solution}
False. First, the $X_i$'s need to be independent. Second, the variance of $\bar X_{n}$ should  be $\sigma^2/n$.
\end{solution}
\end{truefalse}

\begin{truefalse}
  Let $V \sim \chi^2_n$. Claim: For large $n$, $V \stackrel{\cdot}\sim \Norm{n, 2n}$.
\begin{solution}
True. Check the definition.
\end{solution}
\end{truefalse}

\begin{truefalse}
Let $\{X_{n}\}$ be a sequence of rvs with unknown distribution but with finite mean $\mu $ and variance $\sigma^{2}$.
If the average $\bar{X}_{n} \to \Norm{\mu, \sigma/n}$ as $n \to \infty$ then $\bar X_{n}$ must be normally distributed for all $n$, since filling in $ n=1 $ results in $\bar{X}_{n} \sim \Norm{\mu, \sigma^{2}}$
\begin{solution}
It's False. Given a sequence of rvs $\{X_{n}\}$ with finite mean and variance,  $\bar{X}_{n} \to \Norm{\mu, \sigma/n}$ as $n \to \infty$ by the CLT.
\\Also note $X $ instead of $\bar{X}$. X still has the same distribution as it had. The distribution of $\bar{X}_{n}$ converges to $\Norm{\mu, \frac{\sigma^{2}}{n}} $ as $n \to \infty$
\end{solution}
\end{truefalse}


\begin{truefalse}
Let $X_{i} \sim \Pois{\lambda}$, and iid, then $$\left( \frac{ \sum^{n}_{i=1} X_{i}}{\sqrt{n} \sqrt{\lambda}} - \sqrt{n}\sqrt{\lambda} \right) \to \Norm{0,1} \quad \text{as } n \to \infty$$.
\begin{solution}
Correct\\
$$\sqrt{n} \left( \frac{ \bar{X}_{n}  - \lambda}{\sqrt{\lambda}}  \right) $$
$$\sqrt{n} \left( \frac{ \frac{1}{n} \sum^{n}_{i=1} X_{i}  }{\sqrt{\lambda}} - \sqrt{\lambda}  \right) $$
$$ = \left( \frac{ \sum^{n}_{i=1} X_{i}}{\sqrt{n}\lambda} - \sqrt{n}\sqrt{\lambda} \right)$$
$$
\to \Norm{0,1} \quad \text{as } n \to \infty$$
\end{solution}
\end{truefalse}

\begin{truefalse}
Let $T_n = \frac{Z}{\sqrt{V_n/n}}$, with $Z\sim \Norm{0,1}$, $V_n\sim \chi_n^2$, Z and V independent.
Claim: because by SLLN $\frac{V_n}{n}\rightarrow \E{Z_1^2}$ with probability 1, $T_n$ approaches the standard normal distribution.
\begin{solution}
It's True, see BH 480
\end{solution}
\end{truefalse}

\begin{truefalse}
Let $Z_1^2 + \cdots + Z_n^2 = V \sim \chi^2_n$, where $Z_1, \cdots ,Z_n \sim \Norm{0,1}$ are iid,
and given $\E{Z_1^4} = 3$.
Claim: $\V{V} = 2n^2$
\begin{solution}
It's false: $\V{V} = n\V{Z_1^2} = n\left(\E{Z_1^4} - \E{Z_1^2}^2\right) = 2n$.
\end{solution}
\end{truefalse}

\end{document}
