\subsection{BH.9.7}
\label{sec:bh.9.7}


\begin{exercise}
Include your hand written solution for this problem here. Read the instructions of the announcement if you don't know what to do.
\end{exercise}

In the rest of the assignment we use  simulation to numerically solve  BH.9.7. (Note how short  this code is;  amazing, isn't it?)


\begin{minted}[]{python}
import numpy as np
from scipy.stats import uniform
import matplotlib.pyplot as plt

np.random.seed(3)


N = 1000
a, b = 0, 1000000
V = uniform(a, b).rvs(N)

x_range = np.linspace(b / 5, b / 2, num=50)
y = np.zeros_like(x_range)

for i, b in enumerate(x_range):
    payoff = (V - b) * (b >= V / 4)
    y[i] = payoff.mean()


plt.plot(x_range, y)
plt.show()
\end{minted}

\begin{minted}[]{R}
set.seed(3)

N = 1000
a = 0
b = 1000000
V = runif(N, min = a, max = b)

x_range = seq(b / 5, b / 2, length.out = 50)
y = rep(0, length(x_range))


i = 1
for (b in x_range) {
  payoff = (V - b) * (b >= V / 4)
  y[i] = mean(payoff)
  i = i + 1
}

plot(x_range, y, type = "l", col = "blue")
\end{minted}


\begin{exercise}
For the Python code, use the scipy documentation to explain why $V\sim\Unif{[0,10^{6}]}$. For R, explain the function \texttt{runif()}.
\end{exercise}



\begin{exercise}
What are the smallest and the largest value of \verb|x_range|?
\end{exercise}

\begin{exercise}
Run the code above and make a graph. Include the graph in your report, and explain what you see in the graph. For instance, is there a maximum? If so, can you explain where the maximum occurs? Can you explain how the maximum should be?
\end{exercise}


\begin{exercise}
Suppose after seeing the graph of the payoffs, and this graph would only increase, or decrease, how would you change \verb|x_range|? Do you expect to see a maximum?
\end{exercise}




\begin{exercise}
For small N, e.g. N=10, you can get quite strange values. Why is that?
\end{exercise}


\begin{exercise}
Change the acceptance threshold from to $V/4$ to $V/5$ (or $V/6$, or some other value you like), and make a graph of the payoffs.
Include the graph in your report.
\end{exercise}

\begin{exercise}
Change the payoff function to e.g $\sqrt{V-b}$, or some weird function that you like particularly such as $\sin |V-b|$ (any non-trivial function goes).
Make a graph of the  mean and the standard deviation of the payoff. Can you explain your graph?
\end{exercise}
