\section*{Question}
There are three cheese shops in town, shop A, shop B and shop C. $N_1$ customers enter shop A per hour, $N_1\sim \text{Poisson}(\lambda_1)$; $N_2$ customers enter shop B per hour, $N_2\sim \text{Poisson}(\lambda_2)$; $N_3$ customers enter shop C per hour, $N_3\sim \text{Poisson}(\lambda_3)$. $N_1, N_2, N_3$ are independent of each other. Each customer that enters in any of the three shops, buys cheese with probability $p$, independently. Let $X$ be the total number of customers that buys cheese per hour.
% covariance, multivariate normal
\begin{exercise}[1]
 Show that $N_1+N_2$ and $N_2-2N_3$ are \textbf{not} independent of each other.
%Find $\cov{2N_1+N_2, 2N_1-N_2}.$

\begin{solution}
\begin{align*}
   & \cov{N_1+N_2, N_2-2N_3}\\
  = & \cov{N_1,N_2}-\cov{N_1,2N_3}+\cov{N_2,N_2}-\cov{N_2,2N_3}\cdot \cdot \cdot \text{(0.5 point)}\\
  = &  Var(N_2)\\
  =& \lambda_2 > 0.\cdot \cdot \cdot \text{(0.5 point)}
\end{align*}
\end{solution}
\end{exercise}

\begin{exercise}[1]
Let $N=N_1+N_2+N_3$. What is the distribution of  $N$? What is the distribution of $X|N$?  What is the distribution of $X$?
\begin{solution}
Since $N_1, N_2$ and $N_3$ are independent, $N\sim Pois(\lambda_1+\lambda_2+\lambda_3).$(0.5 point)\\
$X|N\sim Bin(N,p),$ and $X\sim Pois((\lambda_1+\lambda_2+\lambda_3) p)$ by the Chicken-egg theory. (0.5 point)
\end{solution}
\end{exercise}


\begin{exercise}[2]
Calculate $\rho_{X,N},$ the correlation between $X$ and $N$.
\begin{solution}
 Let $Y=N-X$ be the number of customers that do not buy cheese. Then we know $Y\sim Pois((\lambda_1+\lambda_2+\lambda_3) q)$ with $q=1-p,$ and $X$ and $Y$ are independent. (0.5 point)
 \begin{align*}
     \cov{N,X}&=\cov{X+Y,X}\\
     &=\cov{X,X}+\cov{Y,X}\\
     &=Var(X)\cdot \cdot \cdot\text{(0.5 point)}\\
     &=(\lambda_1+\lambda_2+\lambda_3)p \cdot \cdot \cdot \text{(0.5 point)}
 \end{align*}
 Then it follows that
 \begin{align*}
     \rho_{X,N}&=\frac{\cov{N,X}}{sd(N)\cdot sd(X)}\\
     &=\frac{(\lambda_1+\lambda_2+\lambda_3)p}{\sqrt{\lambda_1+\lambda_2+\lambda_3}\cdot \sqrt{(\lambda_1+\lambda_2+\lambda_3)p} }\\
     &=\sqrt{p}.\cdot \cdot \cdot\text{(0.5 point)}
 \end{align*}
\end{solution}
\end{exercise}

\noindent Consider the following codes:
\begin{minted}{R}
library(mvtnorm)
set.seed(777)
A<-matrix(c(1,2,2,4),nrow = 2)
B<-c(1,2)
X<-rmvnorm(50,mean=B,sigma=A)
output=cor(X[3,],X[41,])
\end{minted}


\begin{minted}{python}
import random
import numpy as np
random.seed(777)
A = [[1,2],[2,4]]
B = [1,2]
X = np.random.multivariate_normal(B, A, size = 50)
output = np.corrcoef(X[[3,41],:].reshape(4))
\end{minted}

\begin{exercise}[1]
Explain in detail the purpose of \textbf{each line} of the above codes.
\begin{solution}
Line 1: Load the package "mvtnorm" so that we can use the function \textit{rmvnorm}.\\
Line 2: Set a random seed to reproduce the same results.\\
Line 3: Generate a 2$\times$2 identity matrix A=$\begin{pmatrix}1&2\\2&4\end{pmatrix}$.\\
Line 4: Generate a vector $B=(1,2)$.\\
Line 5: Generate 50 multivariate normally distributed variables $\textbf{X}={x_1,x_2}$ with mean equal to $B$ and variance equal to A.\\
Line 6: Calculate the correlation of the 3rd and 41th rows of matrix $X$.\\
\textbf{(0.5 points for mentioning at least 3 of the above.)}
\end{solution}
\end{exercise}
