\section{Question}


Catherine and Denny are playing a game. Each player throws a single (fair) die. The die has three pairs of identical sides: a pair of ones, a pair of twos and a pair of threes. Let $X$ and $Y$ denote the outcome of Catherine and Denny's throw, respectively. The person that throws the highest number wins. If both throw the same number, they have a draw. The final score $C$ of Catherine is determined as follows:
\begin{itemize}
    \item If Catherine loses, then she gets zero points.
    \item If Catherine and Denny draw, then she gets 0.5 point.
    \item If Catherine wins, then her score is the difference $X -Y$ in the numbers they threw.
\end{itemize}
The final score $D$ for Denny is determined analogously. Assume that the dice are fair and that all throws are independent.

\begin{exercise}[1]
Find the joint distribution of $X$ and $Y$ conditional on Catherine winning.
\begin{solution}
The initial distribution is $\P{X = x, Y = y} = 1/9$, $x,y = 1,2,3$. Catherine wins iff $X>Y$, i.e., iff $(X,Y) \in \{(2,1), (3,1), (3,2)\}$. So we have
\begin{align}
    \P{X=x,Y=y | X>Y} &= \frac{\P{X=x,Y=y,X>Y}}{\P{X>Y}} \\
    &= \frac{1/9}{1/3} = 1/3,
\end{align}
for $(x,y) \in \{(2,1), (3,1), (3,2)\}$.
\end{solution}
\end{exercise}

\begin{exercise}[1.5]
What is Catherine's expected score conditional on Catherine winning?
\begin{solution}
We have
\begin{align}
    C = \begin{cases}
    0 &X < Y \\
    1/2 &X = Y \\
    X - Y &X > Y.
    \end{cases}
\end{align}
Hence,
\begin{align}
    \E{C|X>Y} &= \E{X-Y|X>Y} \\
    &= \frac{1}{3} (2-1) + \frac{1}{3} (3-1) + \frac{1}{3} (3-2) \\
    &= 4/3.
\end{align}
\end{solution}
\end{exercise}

\begin{exercise}[1.5]
Determine Catherine's (unconditional) expected score.
\begin{solution}
We have, by the law of total expectation,
\begin{align}
    \E{C} &= \P{X<Y}\E{C|X<Y} + \P{X=Y}\E{C|X=Y} + \P{X>Y}\E{C|X>Y} \\
    &= 0 + \frac{1}{3} \cdot \frac{1}{2} + \frac{1}{3} \frac{4}{3} \\
    &= 11/18.
\end{align}
\end{solution}
\end{exercise}

To spice things up, Catherine and Denny decide to play for money. After playing the dice game and scoring $C$ points, Catherine receives an amount of $S$ euros, where $S$ is determined randomly. Here, conditional on the outcome of $C$, $S$ follows a uniform distribution on $[C,2C]$.
\begin{exercise}[1]
What is the (unconditional) expected reward for Catherine? That is, compute $\E{S}$.
\begin{solution}
By Adam's law,
\begin{align}
    \E{S} &= \E{\E{S|C}} \\
    &= \E{\frac{3}{2}C} \\
    &= \frac{3}{2}\E{C} \\
    &= \frac{3}{2} \frac{11}{18} \\
    &= \frac{33}{36}.
\end{align}
\end{solution}
\end{exercise}