\input{header.tex}

\setcounter{theorem}{34}
\begin{exercise}[BH.2.35]
\begin{solution}~
	\begin{enumerate}
		\item The probability of winning the first game is found by applying the LOTP as follows,
		\begin{align*}
			P(W_{1}) &= P(W_{1}|B)P(B) + P(W_{1}|I)P(I) + P(W_{1}|M)P(M)\\
			& = \frac{9}{10}\cdot\frac{1}{3} + \frac{1}{2}\cdot\frac{1}{3} + \frac{3}{10}\cdot\frac{1}{3}\\
			& = \frac{17}{30}.
		\end{align*}
		\item Denote by $B$, $I$, $M$ the events of having a beginner, intermediate or master opponent, respectively.
		\begin{align*}
			P(W_{2}|W_{1}) &= P(W_{2}|W_{1},B)P(B|W_{1}) + P(W_{2}|W_{1},I)P(I|W_{1}) + P(W_{1}|W_{1},M)P(M|W_{1})\\
			&= P(W_{2}|B)P(B|W_{1}) + P(W_{2}|I)P(I|W_{1}) + P(W_{1}|M)P(M|W_{1})\tag*{(\text{Using that given skill, outcomes are independent})}
		\end{align*}
		Now, using Bayes' rule
		\begin{align*}
			P(B|W_{1})&=\frac{P(W_{1}|B)P(B)}{P(W_{1})}=\frac{9}{17},\\
			P(I|W_{1})&=\frac{P(W_{1}|I)P(I)}{P(W_{1})}=\frac{5}{17},\\
			P(M|W_{1})&=\frac{P(W_{1}|M)P(M)}{P(W_{1})}=\frac{3}{17}.
		\end{align*}
		We can conclude that
		\begin{align*}
			P(W_{2}|W_{1}) &= \frac{9}{10}\frac{9}{17} + \frac{1}{2}\frac{5}{17}+\frac{3}{10}\frac{3}{17}\\
			& = \frac{23}{34}.
		\end{align*}
		Now, look back at your answer to (a). There you found that winning the first game had probability 0.567. Given that you won the first round, you would expect the probability of winning the second round to go up. Indeed, the answer to (b) is 0.676.
		\item Knowing the outcome of one of the matches, we adjust our beliefs on the quality of our opponent. This changes our beliefs on the outcome of the other match. However, if the quality of the opponent is fixed, then outcome of the first match does not affect the outcome of the second match.  
	\end{enumerate}
\end{solution}
\end{exercise}
\input{trailer.tex}
