\input{header.tex}
\setcounter{theorem}{35}
\begin{exercise} BH.8.36 Skip part a since we don't use Jacobians in this year (i.e., 2022-2023).
The hints and solutions still contain the answers for part a for completeness sake, but you should
skip this, unless you are interested in learning more.

\begin{hint}
a. See BH.8.5.1. The exponential is a special case of the gamma distribution. See also BH.8.34.c. $T_1/T_2$ is a function of $T_1/(T_1+T_2)$.

b. This can be solved with a joint distribution function and integration over the event $\{T_1<T_2\}$. However, we can use Exercise BH.7.10 or BH.7.1.24.

c. First she has to wait for the first server to become free. This is the minimum of the two exponentials.
With $\P{T_1<T_2}$ server 1 is the first.
What is the probability that the other server is empty first? Then, once she is at a server, what is her expected service time? The total time in the system is the time in queue plus the service time.
\end{hint}
\begin{solution}
a. This year (2022-2023) you are allowed to skip part 1. It's included for the sake of completeness.

I did not attempt any smart tricks. Take as transform $u=s/t$ and $v=s+t$, where I associate $s$ to $T_{1}$ and $t$  to $T_2$. Then,
\begin{align*}
\frac{\partial (u,v)}{\partial (s,t)} &=
  \begin{pmatrix}
    1/t & -s/t^{2} \\
1 & 1
  \end{pmatrix} = \frac 1 t + \frac{s}{t^{2}} = \frac{t+s}{t^{2}} = \frac{v}{t^{2}}
\end{align*}
With a bit of algebra: $s=uv/(u+1)$ and $t=v/(u+1)$.
Therefore the Jacobian we need becomes
\begin{equation}
\label{eq:12}
\frac{\partial (s,t)}{\partial (u,v)} = \frac{t^{2}}{v} = \frac{v^{2}}{(u+1)^2} \frac{1}{v} = \frac{v}{(u+1)^{2}}.
\end{equation}
Next,
\begin{align*}
  f_{U, V}(u,v) &= f_{T_1,T_2}(s,t) \frac{\partial (s,t)}{\partial (u,v)}
  =  f_{T_1}(uv/(u+1) f_{T_2}(v/(u+1)) \frac{v}{(u+1)^{2}} \\
  &=  \lambda^{2} e^{-\lambda uv/(u+1)} e^{- \lambda v/(u+1)} \frac v{(u+1)^{2}} \\
  &=  \lambda^{2} \exp(-\lambda v) v \cdot \frac 1{(u+1)^{2}}.
\end{align*}
Clearly, this is a product of a function $f_{V}(f)$ that depends only  $v$ and another function $f_U(u)$ that depends  only $u$. Hence, $U$ and $V$ are independent.

b. With the hint,
\begin{align*}
\P{T_1<T_2}
&= \int_0^{\infty} \P{T_1< T_2\given T_1=s} f_{T_1}(s)\d s \\
&= \int_0^{\infty} \P{s< T_2\given T_1=s} \lambda_1e^{-\lambda_1 s} \d s \\
&= \lambda_{1} \int_0^{\infty} e^{-\lambda_{2} s} e^{-\lambda_1 s} \d s = \frac{\lambda_1}{\lambda_1+\lambda_2}.
\end{align*}

c. See the hint. Alice first has to wait for the first server to become free. The expected time in queue is $1/(\lambda_1+\lambda_2)$. If server 1 is the first, then Alice spends a  time $1/\lambda_1$ in service. Thus, the total time is
\begin{align*}
  \frac{1}{\lambda_1+\lambda_2} +
  \frac{\lambda_1}{\lambda_1+\lambda_2} \frac{1}{\lambda_{1}} +
  \frac{\lambda_2}{\lambda_1+\lambda_2} \frac{1}{\lambda_{2}}  =
\frac{3}{\lambda_1+\lambda_{2}}.
\end{align*}
\end{solution}
\end{exercise}
\input{trailer.tex}
